---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# tidyhte

<!-- badges: start -->
[![Codecov test coverage](https://codecov.io/gh/ddimmery/tidyhte/branch/main/graph/badge.svg)](https://codecov.io/gh/ddimmery/tidyhte?branch=main)
[![R-CMD-check](https://github.com/ddimmery/tidyhte/workflows/R-CMD-check/badge.svg)](https://github.com/ddimmery/tidyhte/actions)
<!-- badges: end -->

The goal of tidyhte is to ...

## Installation

You can install the released version of tidyhte from [CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("tidyhte")
```

And the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("ddimmery/tidyhte")
```

You'll still need to render `README.Rmd` regularly, to keep `README.md` up-to-date. `devtools::build_readme()` is handy for this. You could also use GitHub Actions to re-render `README.Rmd` every time you push. An example workflow can be found here: <https://github.com/r-lib/actions/tree/master/examples>.

```{r demo}
rnorm(10)
```

