---
title: Example HTE Analysis
author: Drew Dimmery (drew.dimmery@gmail.com)
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
    html_document:
        toc: true
        toc_float: true
        toc_depth: 3
        theme: flatly
    rmarkdown::html_vignette:
vignette: >
  %\VignetteIndexEntry{Example HTE Analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<style type="text/css">
div.main-container {
  max-width: 1800px;
  margin-left: auto;
  margin-right: auto;
}
</style>

# Introduction

In this example analysis, I will demonstrate how to run an analysis of heterogeneous treatment effects in a randomized control trial using the methods of Kennedy (2020). This implies that propensity scores are _known_, so estimates will generally be unbiased.

This analysis will focus on using machine learning ensembles (through SuperLearner) to estimate nuisance functions and then provide a tibble of estimates of conditional treatment effects along with their associated standard errors.

# Load packages and setup environment

```{r setup, print=FALSE, message=FALSE}
# library(tidyhte)
devtools::load_all()
library(ggplot2)
library(dplyr)
# library(alethea)
# library(extrafont)
# loadfonts()

OUTPUT_FOLDER <- "~"
```


# Simulate data

If real data is used, simply replace this block with an appropriate `readr::read_csv` call or equivalent, creating a tibble. I will assume this tibble is stored as `data` for the remainder of this document.

Note that datatypes can be either continuous or discrete, and that there can be columns in the tibble that are not included in any resulting anayses.

```{r sim_data}
set.seed(100)
n <- 5000
data <- tibble(
    uid = 1:n
) %>%
    mutate(
        a = rbinom(n, 1, 0.3),
        ps = rep(0.3, n),
        x1 = rnorm(n),
        x2 = factor(sample(1:4, n, prob = c(1 / 100, 39 / 100, 1 / 5, 2 / 5), replace = TRUE)),
        x3 = factor(sample(1:3, n, prob = c(1 / 5, 1 / 5, 3 / 5), replace = TRUE)),
        # x4 = (x1 + rnorm(n)) / 2,
        x4 = runif(n),
        x5 = rnorm(n),
        y = (
            a + x1 - a * (x1 - mean(x1)) + (4 * rbinom(n, 1, 0.5) - 1) * a * (x2 == 2) +
            a * (x2 == 3) + 0.5 * a * (x2 == 4) +
            rnorm(n)
        ),
        w = 0.1 + rexp(n, 1 / 0.9)
    )
```

# Define variables

The primary things that require user-specification here are the covariate names.

```{r define}
# Define the (unquoted) username column
userid <- rlang::expr(uid)
weights <- rlang::expr(w)
# Define the (quoted) variable name of the propensity score
propensity_score_variable_name <- "ps"
# Define the (quoted) names of continuously-valued (i.e. double) covariates
continuous_covariates <- c("x1")
# Define the (quoted) names of discrete-valued (i.e. factor) covariates
discrete_covariates <- c("x2", "x3")
# Define the moderators to measure the MCATE for
continuous_moderators <- rlang::exprs() # x1, x4, x5)
discrete_moderators <- rlang::exprs(x2, x3)
moderators <- c(continuous_moderators, discrete_moderators)
# Uncomment the following to just use the union of all covariates above
# moderators <- rlang::syms(c(continuous_covariates, discrete_covariates))
# Names of all underlying covariates
model_covariate_names <- c(continuous_covariates, discrete_covariates)
model_covariates <- rlang::syms(model_covariate_names)

# Replace y with the (unquoted) name of the outcome variable of interest.
outcome_variable <- rlang::expr(y)
# Replace a with the (unquoted) name of the treatment variable of interest.
treatment_variable <- rlang::expr(a)
```

# Define Propensity Score Model

Since in this example we know the exact propensity scores for all units, we use the `Known_cfg` config type.

```{r pscore_def}
trt.cfg <- Known_cfg$new(propensity_score_variable_name)
```

# Define Outcome Model

We estimate the outcome (T-learner) plugin estimate using an ensemble of machine learning models, including a wide array of model complexities from linear models, GAMs, regularized regressions, random forests, and gradient-boosted decision trees.

Each individual component of the model provides a list of hyperparameters, over which a full cross-product is taken and all resulting models are estimate. For instance, `SL.glmnet` sweeps over two hyperparameters (two losses and four mixing parameters between ridge and Lasso). This implies that there will be 2 times 4 models estimated and incorporated into the ensemble. Note that `SL.glmnet` automatically tunes the regularization parameter using `cv.glmnet`, so this is not included as a hyperparameter.

To run the full set of models, uncomment the lines below.
```{r outcome_def}
regression.cfg <- SLEnsemble_cfg$new(
  learner_cfgs = list(
    SLLearner_cfg$new(
      "SL.glm"
    ),
    SLLearner_cfg$new(
      "SL.glm.interaction"
    ),
    SLLearner_cfg$new(
      "SL.mean"
    ),
    SLLearner_cfg$new(
      "SL.glmnet",
      list(
        alpha = c(0.05, 0.15, 0.2, 0.25, 0.5, 0.75),
        loss = c("deviance")
      )
    ),
    SLLearner_cfg$new(
      "SL.glmnet.interaction",
      list(
        alpha = c(0.05, 0.15, 0.2, 0.25, 0.5, 0.75),
        loss = c("deviance")
      )
    )
    # SLLearner_cfg$new(
    #   "SL.ranger",
    #   list(
    #     num.trees = c(25, 50, 100, 250),
    #     splitrule = c("gini")
    #   )
    # ),
    # SLLearner_cfg$new(
    #   "SL.xgboost",
    #   list(
    #     ntrees = c(50, 100, 250, 500),
    #     max_depth = c(1, 3, 5),
    #     shrinkage = c(0.1)
    #   )
    # )
  )
)
```


# Define Quantities of interest

Quantities of Interest determine how results are reported to the user. You can think about this as determining, for instance how results should be plotted in a resulting chart.

For simplicity, this example simply provides results in one of two ways:
- Discrete covariates are stratified and the conditional effect is plotted at each distinct level of the covariate.
- Continuous covariates have the effect surface estimated using local-linear regression via `nprobust` of Calonico, Cattaneo and Farrell (2018). See, similarly, Kennedy, Ma, McHugh and Small (2017) for justification of this approach. Results are obtained for a grid of 100 quantiles across the domain of the covariate.

An additional quantity of interest provided is the variable importance of a learned joint model of conditional effects (over all covariates). The approach implemented is described in Williamson, Gilbert, Carone and Simon (2020).

```{r def_qoi}
qoi.list <- list()
for (cov in continuous_moderators) {
    qoi.list[[rlang::as_string(cov)]] <- KernelSmooth_cfg$new(neval = 100)
}
for (cov in discrete_moderators) {
    qoi.list[[rlang::as_string(cov)]] <- Stratified_cfg$new(cov)
}

qoi.cfg <- QoI_cfg$new(
    mcate = MCATE_cfg$new(cfgs = qoi.list),
    # Don't use! The PCATE is still broken.
    pcate = PCATE_cfg$new(
        cfgs = qoi.list,
        effect_cfg = regression.cfg,
        model_covariates = model_covariate_names,
        # In a real analysis, this should be higher so as to reduce Monte Carlo error.
        # num_mc_samples = list(x1 = 25, x2 = 1000, x3 = 1000, x4 = 25, x5 = 25)
        num_mc_samples = list(x1 = 5, x2 = 10, x3 = 10, x4 = 5, x5 = 5)
    ),
    vimp = VIMP_cfg$new(model_cfg = regression.cfg),
    diag = Diagnostics_cfg$new(
        outcome = c("SL_risk", "SL_coefs", "MSE")
        #effect = c("SL_risk", "SL_coefs")
    )
)
```

## Combine all configs
```{r}
cfg <- HTE_cfg$new(
    treatment = trt.cfg,
    outcome = regression.cfg,
    qoi = qoi.cfg
)
```

## Future API

The above configuration files are not very user-friendly. A more tidy-friendly API will follow the lead of various packages which use "recipes" to define how models will be trained and executed.

Soon it will look more like the following:

```{r eval=FALSE}
initialize_config() %>%
    add_propensity_model("known", pscore) %>%
    add_outcome_model("glmnet", interactions = FALSE, alpha = c(0.05, 0.15, 0.2, 0.25)) %>%
    add_outcome_model("glmnet", interactions = TRUE, alpha = c(0.05, 0.15, 0.2, 0.25)) %>%
    add_outcome_model("xgboost", ntrees = c(50, 100, 250, 500), shrinkage = c(0.01, 0.1)) %>%
    add_quantity_of_interest("MCATE", x1, "KernelSmooth") %>%
    add_quantity_of_interest("MCATE", x2, "Stratified") %>%
    add_quantity_of_interest("VIMP", x1, x2) %>%
    add_propensity_diagnostic("AUC") %>%
    add_propensity_diagnostic("SL coefficients") %>%
    add_outcome_diagnostic("MSE") %>%
    add_outcome_diagnostic("SL coefficients")
```

# Estimate Models

To actually perform the estimation, the following will be sufficient. Note that the configuration of covariate names at the top of the document makes all of this a little more complex with all the curly-brackets and bangs.

```{r estimate, message=FALSE}
data %>%
    attach_config(cfg) %>%
    make_splits({{ userid }}, .num_splits = 4) %>%
    produce_plugin_estimates(
        {{ outcome_variable }},
        {{ treatment_variable }},
        !!!model_covariates,
        .weights = {{ weights }}  # nprobust does not currently support weights
    ) -> prepped_data

construct_pseudo_outcomes(prepped_data, {{ outcome_variable }}, {{ treatment_variable }}) %>%
    estimate_QoI(!!!moderators) -> results
```


```{r show_qoi, message=FALSE}
results
```


# ATEs 
```{r ates}
filter(results, grepl("SATE|PATE", estimand))
```

```{r joint}
fm <- paste0(".pseudo_outcome~", paste(purrr::map_chr(moderators, rlang::as_string), collapse = "+"))

df_jm <- construct_pseudo_outcomes(prepped_data, {{ outcome_variable }}, {{ treatment_variable }})
simple_joint_model <- stats::lm(as.formula(fm), df_jm)

print(summary(simple_joint_model))
```


# Write results to file 

```{r write, eval=FALSE}
readr::write_csv(results_dr, file.path(OUTPUT_FOLDER, paste0("results_", rlang::as_string(outcome_variable), ".csv.gz")))
```

# Plots
## Plot Ensemble Coefficients

```{r sl_coef}
plot_list <- list()
filter(results, grepl("SL coefficient", estimand)) %>%
mutate(level = factor(level, levels = c("Control Response", "Treatment Response", "Effect Surface"))) %>%
ggplot(aes(
            x = reorder(term, estimate),
            y = estimate,
            ymin = estimate - 1.96 * std_error,
            ymax = estimate + 1.96 * std_error
    )) +
    geom_abline(intercept = 0, slope = 0, linetype = "dashed") +
    geom_pointrange() +
    expand_limits(y = 0) +
    scale_x_discrete("Model name") +
    scale_y_continuous("Coefficient in SuperLearner Ensemble") +
    facet_wrap(~level) +
    coord_flip() +
    ggtitle("SuperLearner Ensemble") +
    theme_minimal() -> gp
plot_list <- c(plot_list, list(gp))
print(gp)
```

## Plot risk for each submodel
```{r sl_risk}
filter(results, grepl("SL risk", estimand)) %>%
mutate(level = factor(level, levels = c("Control Response", "Treatment Response", "Effect Surface"))) %>%
ggplot() +
    geom_abline(intercept = 0, slope = 0, linetype = "dashed") +
    geom_pointrange(
        aes(
            x = reorder(term, -estimate),
            y = estimate,
            ymin = estimate - 1.96 * std_error,
            ymax = estimate + 1.96 * std_error)
        ) +
    expand_limits(y = 0) +
    scale_x_discrete("Model name") +
    scale_y_continuous("CV Risk in SuperLearner Ensemble") +
    facet_wrap(~level, scales = "free_x") +
    coord_flip() +
    ggtitle("Submodel Risk Estimates") +
    theme_minimal() -> gp
plot_list <- c(plot_list, list(gp))
print(gp)
```

## Plot VIMP

```{r vimp}
ggplot(filter(results, estimand == "VIMP")) +
    geom_abline(intercept = 0, slope = 0, linetype = "dashed") +
    geom_pointrange(
        aes(x = term, y = estimate, ymin = estimate - 1.96 * std_error, ymax = estimate + 1.96 * std_error)) +
    expand_limits(y = 0) +
    scale_x_discrete("Covariate") +
    scale_y_continuous("Reduction in R² from full model") +
    coord_flip() +
    ggtitle("Covariate Importance") +
    theme_minimal() -> gp
plot_list <- c(plot_list, list(gp))
print(gp)
```

## Plot Continuous Covariates' MCATE
```{r cts_mcate_plot, message=FALSE}
for (cov in continuous_moderators) {
    ggplot(filter(results, estimand == "MCATE", term == cov)) +
        geom_abline(intercept = 0, slope = 0, linetype = "dashed") +
        geom_ribbon(
            aes(x = value, ymin = estimate - 1.96 * std_error, ymax = estimate + 1.96 * std_error), alpha = 0.75
        ) +
        geom_line(
            aes(x = value, y = estimate)
        ) +
        expand_limits(y = 0) +
        scale_x_continuous("Covariate level") +
        scale_y_continuous("CATE") +
        ggtitle(paste("Marginal effects across", cov)) +
        theme_minimal() -> gp
    plot_list <- c(plot_list, list(gp))
    print(gp)
}
```

## Plot Continuous Covariates' PCATE
```{r cts_pcate_plot, message=FALSE}
for (cov in continuous_moderators) {
    ggplot(filter(results, estimand == "PCATE", term == cov)) +
        geom_abline(intercept = 0, slope = 0, linetype = "dashed") +
        geom_ribbon(
            aes(x = value, ymin = estimate - 1.96 * std_error, ymax = estimate + 1.96 * std_error), alpha = 0.75
        ) +
        geom_line(
            aes(x = value, y = estimate)
        ) +
        expand_limits(y = 0) +
        scale_x_continuous("Covariate level") +
        scale_y_continuous("CATE") +
        ggtitle(paste("Partial effects across", cov)) +
        theme_minimal() -> gp
    plot_list <- c(plot_list, list(gp))
    print(gp)
}
```

## Plot Discrete Covariates' MCATE
```{r discrete_mcate_plot}
for (cov in discrete_moderators) {
    ggplot(filter(results, estimand == "MCATE", term == cov)) +
        geom_abline(intercept = 0, slope = 0, linetype = "dashed") +
        geom_pointrange(
            aes(x = level, y = estimate, ymin = estimate - 1.96 * std_error, ymax = estimate + 1.96 * std_error)) +
        expand_limits(y = 0) +
        scale_x_discrete("Covariate level") +
        scale_y_continuous("CATE") +
        ggtitle(paste("Marginal effects across", cov)) +
        theme_minimal() -> gp
    plot_list <- c(plot_list, list(gp))
    print(gp)
}
```

## Plot Discrete Covariates' PCATE
```{r discrete_pcate_plot}
for (cov in discrete_moderators) {
    ggplot(filter(results, estimand == "PCATE", term == cov)) +
        geom_abline(intercept = 0, slope = 0, linetype = "dashed") +
        geom_pointrange(
            aes(x = level, y = estimate, ymin = estimate - 1.96 * std_error, ymax = estimate + 1.96 * std_error)) +
        expand_limits(y = 0) +
        scale_x_discrete("Covariate level") +
        scale_y_continuous("CATE") +
        ggtitle(paste("Partial effects across", cov)) +
        theme_minimal() -> gp
    plot_list <- c(plot_list, list(gp))
    print(gp)
}
```

```{r save_plots, eval=FALSE}
# Might need to embed fonts depending on the theme.
plot_filename <- file.path(OUTPUT_FOLDER, paste0("plots_", rlang::as_string(outcome_variable), ".pdf"))
pdf(file = plot_filename, height = 7, width = 7, fonts = "Arial Narrow")
plot_list
dev.off()
embed_fonts(normalizePath(plot_filename))
```


# Session Info
```{r session_info}
print(sessionInfo())
```